\documentclass[11pt, letterpaper]{article}
\usepackage[margin=2cm]{geometry}
\usepackage{fourier}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{placeins}
\usepackage{bookmark}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{parskip}
\usepackage{epstopdf}
\usepackage[sorting=none,style=ieee]{biblatex}
\DeclareUnicodeCharacter{2009}{\,}
\setcounter{secnumdepth}{-\maxdimen}

\title{cuTimeWarp: Accelerating Soft Dynamic Time Warping on GPU}

\author{Alex Kyllo \and Afrooz Rahmati}

\addbibresource{cuTimeWarp.bib}

\begin{document}

\maketitle

\begin{abstract}

This report explores techniques for optimizing the computation of Soft Dynamic
Time Warping, a differentiable sequence dissimilarity measure, on graphics
processing units (GPU), for the purpose of enabling high-performance machine
learning on time series datasets.

\end{abstract}

\section{Introduction}

Time series machine learning is a research area with countless useful
applications such as recognizing sounds and gestures. Clustering or classifying
large time series datasets is challenging partly because of the need to define a
measure of dissimilarity between two time series. Practical applications require
finding common structure despite different speeds or phases; a word means the
same whether spoken quickly or slowly. Another requirement is that the measure
must be differentiable so that its gradient can be used as a loss function to
minimize in model fitting. Finally, the measure must be efficient to calculate.
To this end we will explore GPU acceleration of Soft Dynamic Time Warping
(Soft-DTW) \cite{cuturi_soft-dtw_2018}, a differentiable dissimilarity measure,
to enable high performance time series machine learning.

Dynamic Time Warping (DTW) is an algorithm to compute the dissimilarity between
two time series, which may vary in speed and phase. The basic algorithm for DTW
is to use Bellman’s recursion, a dynamic programming technique, to find the
lowest-cost path diagonally across a pairwise distance matrix. The computation
cost for this approach is quadratic(mn) for time series vectors of length m and
n \cite{cuturi_soft-dtw_2018}. The formula for the DTW between time series x and
y is: 

$$DTW(x,y) = min_{\pi}\sqrt{\sum_{(i,j)\in\pi}d(x_{i},y_{j})^2}$$

Where $d(x_i,y_j)^2$ is the cost function, typically pairwise squared Euclidean
distance. The loss function for DTW is not differentiable due to the min
operation within the formula; a small change in the input time series may result
in zero change in the path cost. However, we can create a differentiable version
called Soft-DTW by replacing the min with a soft-min function
\cite{cuturi_soft-dtw_2018}:

$$\text{soft-min}_\gamma(a_1,...,a_n) = -\gamma log\sum_{i}e^{-a_i/\gamma}$$

Hence, Soft-DTW is parameterized by the smoothing constant gamma, which becomes
a tunable hyperparameter in machine learning model training applications.

A common technique in machine learning with Soft-DTW is the computation of
barycenters, which are centroids within the space of a set of time series. The
differentiability of Soft-DTW allows for barycenter finding via gradient
descent, and then new observations can be clustered or classified by finding the
closest barycenter.

\section{Methods}

\section{Results}

\section{Discussion}

\printbibliography[]
\end{document}